Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis in an Idealized, Axisymmetric Framework

### 2018-07-09 paper.tex
Correct typo in abstract.
Add Emanuel and Rotunno (2011) citation and correct parenthetical to reflect this. Correct typo ('warm care').

After undergoing a remarkably rapid intensification (RI), Hurricane Patricia (2015) set a new record as the strongest tropical cyclone (TC) ever observed in the Western Hemisphere (\citeauthor{Kimberlainetal2016} \citeyear{Kimberlainetal2016}; \citeauthor{Rogersetal2017} \citeyear{Rogersetal2017}).
%High-altitude dropsonde observations taken during the Tropical Cyclone Intensity (TCI) experiment captured this RI in unprecedented detail \citep{DoyleTCI}.
TCI dropsonde observations collected during this RI period revealed dramatic changes in the cold-point tropopause height and upper-level static stability \citep{DuranMolinari2018}.
In particular, when Patricia was at tropical storm intensity shortly before RI commenced, a strong inversion layer existed just above the cold-point tropopause.
During the first half of the RI period, this inversion layer weakened throughout Patricia's inner core, with the weakening most pronounced over the developing eye.
By the time the storm reached its maximum intensity of 95 m s\textsuperscript{-1}, the inversion layer over the eye had disappeared almost completely, which was accompanied by a greater than 1-km increase in the tropopause height.
Meanwhile over the eyewall region, the static stability increased and the tropopause remained near its initial level.
%The mechanisms that might have led to this tropopause-layer variability will be investigated in the current paper using idealized simulations.

@@

An idealized simulation of a TC analyzed by \cite{OhnoSatoh2015} suggested that the development of an upper-level warm core near the 13-km level acted to decrease the static stability near the tropopause within the eye.
During the early stage of development in their simulation, large static stability existed above 16 km at all radii (their Fig. 9c).
However, after the storm's intensification, the static stability within the eye above 16 km was markedly smaller (their Fig. 10c).
Although the mechanisms that might drive this static stability evolution have not been examined explicitly, it might be related to the development of an upper-tropospheric warm care within the eye.

@@

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009}.
The equations of motion were integrated on a 3000-km-wide, 30-km-deep axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
The computations were performed on an \textit{f}-plane at 15\textdegree{N} latitude, over a sea surface with constant temperature of 30.5\textdegree C, which matches that observed near Hurricane Patricia (2015; \citeauthor{Kimberlainetal2016} \citeyear{Kimberlainetal2016}).
Horizontal turbulence was parameterized using the Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, pg. 1773), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} scheme, and radiative heating tendencies were computed every two minutes using the Rapid Radiative Transfer Model for GCMs (RRTMG) longwave and shortwave schemes \citep{Iacono}.
The initial temperature and humidity field was horizontally homogeneous and determined by averaging all Climate Forecast System Reanalysis (CFSR) grid points within 100 km of Patricia's center of circulation at 18 UTC 21 October 2015.
%A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
%Above 19 km, where few TCI observations were available, the temperature profile was taken from the Climate Forecast System Reanalysis (CFSR) grid point nearest Patricia's storm center, valid at 18 UTC 21 October, 2015.
%Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, relative humidity was set equal to 50\% above 11.5 km (the level above which temperature dropped below -40\textdegree C).
The vortex described in \citeauthor{RotunnoEmanuel} (\citeyear{RotunnoEmanuel}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

Two distinct maxima of vertical eddy diffusivity developed in the tropopause layer as the storm intensified (Fig.~\ref{fig:diff}).
Comparison of these turbulent regions to the $N^2$ tendencies in Figs. \ref{fig:stab-24-48}c and \ref{fig:stab-48-72}c reveals that the layers in which vertical eddy diffusivity maximized corresponded to layers of destabilization due to vertical turbulence.
Just outside of these layers, however, vertical turbulence acted to increase $N^2$.
The large vertical gradient of vertical eddy diffusivity near the tropopause played an important role in developing the lower-stratospheric stable layer during RI.
These results support the hypothesized role of turbulence in setting the outflow-layer $\theta$ stratification in \cite{RotunnoEmanuel}.

>>>>

After undergoing a remarkably rapid intensification (RI), Hurricane Patricia (2015) attained the strongest wind speed ever recorded in a tropical cyclone (TC) (\citeauthor{Kimberlainetal2016} \citeyear{Kimberlainetal2016}; \citeauthor{Rogersetal2017} \citeyear{Rogersetal2017}).
%High-altitude dropsonde observations taken during the Tropical Cyclone Intensity (TCI) experiment captured this RI in unprecedented detail \citep{DoyleTCI}.
TCI dropsonde observations collected during this RI period revealed dramatic changes in the cold-point tropopause height and upper-level static stability \citep{DuranMolinari2018}.
In particular, when Patricia was at tropical storm intensity shortly before RI commenced, a strong inversion layer existed just above the cold-point tropopause.
During the first half of the RI period, this inversion layer weakened throughout Patricia's inner core, with the weakening most pronounced over the developing eye.
By the time the storm reached its maximum intensity of 95 m s\textsuperscript{-1}, the inversion layer over the eye had disappeared almost completely, which was accompanied by a greater than 1-km increase in the tropopause height.
Meanwhile over the eyewall region, the static stability increased and the tropopause remained near its initial level.
%The mechanisms that might have led to this tropopause-layer variability will be investigated in the current paper using idealized simulations.

@@

An idealized simulation of a TC analyzed by \cite{OhnoSatoh2015} suggested that the development of an upper-level warm core near the 13-km level acted to decrease the static stability near the tropopause within the eye.
During the early stage of development in their simulation, large static stability existed above 16 km at all radii (their Fig. 9c).
However, after the storm's intensification, the static stability within the eye above 16 km was markedly smaller (their Fig. 10c).
Although the mechanisms that might drive this static stability evolution have not been examined explicitly, it might be related to the development of an upper-tropospheric warm core within the eye.

@@

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009}.
The equations of motion were integrated on a 3000-km-wide, 30-km-deep axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
The computations were performed on an \textit{f}-plane at 15\textdegree{N} latitude, over a sea surface with constant temperature of 30.5\textdegree C, which matches that observed near Hurricane Patricia (2015; \citeauthor{Kimberlainetal2016} \citeyear{Kimberlainetal2016}).
Horizontal turbulence was parameterized using the Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, pg. 1773), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} scheme, and radiative heating tendencies were computed every two minutes using the Rapid Radiative Transfer Model for GCMs (RRTMG) longwave and shortwave schemes \citep{Iacono}.
The initial temperature and humidity field was horizontally homogeneous and determined by averaging all Climate Forecast System Reanalysis (CFSR) grid points within 100 km of Patricia's center of circulation at 18 UTC 21 October 2015.
%A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
%Above 19 km, where few TCI observations were available, the temperature profile was taken from the Climate Forecast System Reanalysis (CFSR) grid point nearest Patricia's storm center, valid at 18 UTC 21 October, 2015.
%Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, relative humidity was set equal to 50\% above 11.5 km (the level above which temperature dropped below -40\textdegree C).
The vortex described in \citeauthor{RotunnoEmanuel1987} (\citeyear{RotunnoEmanuel1987}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

Two distinct maxima of vertical eddy diffusivity developed in the tropopause layer as the storm intensified (Fig.~\ref{fig:diff}).
Comparison of these turbulent regions to the $N^2$ tendencies in Figs. \ref{fig:stab-24-48}c and \ref{fig:stab-48-72}c reveals that the layers in which vertical eddy diffusivity maximized corresponded to layers of destabilization due to vertical turbulence.
Just outside of these layers, however, vertical turbulence acted to increase $N^2$.
The large vertical gradient of vertical eddy diffusivity near the tropopause played an important role in developing the lower-stratospheric stable layer during RI.
These results support the hypothesized role of turbulence in setting the outflow-layer $\theta$ stratification in \cite{EmanuelRotunno2011}.
